To solve the compatibility condition for the **COSMOS** model, we define the precise relationship between global stroboscopic evolution and its local projection. 

Here is the LaTeX subsection for **Section 8.2.1**:

```latex
% =====================================================================================
\subsection*{8.2.1 Precise Compatibility \& Stability Thresholds}
% =====================================================================================

\begin{FormalBox}{The Commutativity Constraint}
Let $\mathcal{B}(\Substrate)$ be the algebra of bounded operators on the substrate. We define the \textbf{Stability Error} $\epsilon$ as the operator norm of the mismatch between the global update and the induced local dynamics:
\[
\epsilon \coloneqq \left\| \Proj \Floquet - \Floquet_{\mathrm{loc}} \Proj \right\|_{\mathrm{op}} = \sup_{\ket{\psi} \in \Substrate, \bra{\psi}\ket{\psi}=1} \left| (\Proj \Floquet - \Floquet_{\mathrm{loc}} \Proj) \ket{\psi} \right|
\]
The emergent sector $\Substrate_{\mathrm{loc}}$ is \textbf{stably coherent} if there exists a local propagator $\Floquet_{\mathrm{loc}}$ such that:
\[
\epsilon < \delta_{\mathrm{coh}}
\]
where $\delta_{\mathrm{coh}}$ is the \emph{coherence margin} defined by the integration operator $\Integr$ (typically related to the spectral gap of the global generator $\Htot$).
\end{FormalBox}

\begin{MeaningBox}{Dynamical Drift and Decoherence}
The condition $\epsilon \approx 0$ implies that the local "map" is a faithful representation of the global "land." If $\epsilon$ exceeds the threshold $\delta_{\mathrm{coh}}$, the local sector undergoes \textbf{representation drift}. 

In this state, the effective operator $\Emerg$ can no longer consistently organize histories; the projected sector "decouples" from the global rhythm. This is the origin of decoherence: not necessarily an interaction with an external environment (as there is no "outside" the substrate), but a failure of the local projection to track the global periodic update.
\end{MeaningBox}

\begin{ChildBox}{The Movie Projector}
Imagine a movie projector ($\Floquet$) spinning a film. The projection operator ($\Proj$) is the lens that puts the light on the screen. If the lens and the film-strip are perfectly synchronized, the movie looks sharp ($\epsilon < \delta_{\mathrm{coh}}$). But if the lens starts wobbling or the film slips, the picture gets blurry and eventually disappears into static. That "blurriness" is $\epsilon$.
\end{ChildBox}

\begin{TestBox}{Phase Transitions in Time Crystals}
What would count as evidence? A sharp transition in the "classicality" of a system. If we tune a system's parameters such that $\epsilon$ crosses $\delta_{\mathrm{coh}}$, we should observe a sudden loss of periodic recurrence—a "melting" of the emergent time crystal into a state of maximum entropy (thermalization) where $\Emerg$ no longer yields stable geometry.
\end{TestBox}
```
