\documentclass[11pt]{article}

% Packages
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{geometry}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{cite}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{titling}

\geometry{
  a4paper,
  margin=0.8in,
  top=1in
}
\begin{document}
% Move title closer to the top
\setlength{\droptitle}{-5em}  % try -1em, -2em, -3em etc.
\setlength{\skip\footins}{2em}

\title{\textbf{The Time Crystalline Holographic Universe}\\[0.5em]
A Non-Dual Model of Conciousness}

\author{Julien Xavier Steff}
\affil[1]{Science Coherence Institute}
\date{} % no date

\maketitle

\begin{abstract}
We propose a unified physical framework in which the Universe is described as a self-referential quantum-informational holographic process unfolding on a time-crystalline substrate. The framework integrates principles from quantum information theory, condensed-matter physics, cosmology, and quantum biology to examine how a spontaneously oscillatory, non-equilibrium phase of matter could serve as the fundamental fabric of reality. Within this substrate intrinsic Floquet-like update cycles provide the fundamental temporal ordering, generating stable patterns of coherence without requiring an external time parameter. 
We show conceptually how such a system could give rise to self-referential holographic conciousness, spontaneous symmetry breaking of
global coherence, and the emergence of multiple observer-subsystems. 
We formulate the substrate’s self-referential evolution law, analyze non-equilibrium coherence instabilities using synchronization theory, and derive a free-energy–minimizing correction dynamic that restores global phase alignment.
We connect these mechanisms to quantum-biological coherence and outline observational consequence, produces testable predictions and unifies quantum information, non-equilibrium physics, biology and cosmology under a single feedback principle.
Within this architecture, gravitational behavior emerges from systematic shifts in coherence produced by matter, quantum measurement corresponds to local stabilization events in the substrate, and biological systems appear as long-lived high-coherence domains. The framework eliminates singularities by enforcing finite-density Planck cores and yields concrete observational signatures across mesoscopic, astrophysical, and cosmological scales.
The result is a coherent, testable, and mathematically consistent unification of quantum information, non-equilibrium physics, gravitation, biology, and cosmology within a single self-updating time-crystalline holographic substrate.
\end{abstract}

\input{Section_1}
\input{Section_2}
\input{Section_3}
\input{Section_4}

\bibliographystyle{plain}
\begin{thebibliography}{9}
\bibitem{wheeler1990information} Wheeler, J. A. (1990). \emph{Information, physics, quantum: The search for links.} In Complexity, Entropy, and the Physics of Information (Zurek, W. H., Ed.). Addison-Wesley.
\bibitem{lloyd2002computational} Lloyd, S. (2002). Computational capacity of the universe. \emph{Physical Review Letters}, 88(23), 237901.
\bibitem{wolfram2020models} Wolfram, S. (2020). A Class of Models with the Potential to Represent Fundamental Physics. \emph{Wolfram Physics Project}.
\bibitem{wilczek2012quantum} Wilczek, F. (2012). Quantum Time Crystals. \emph{Physical Review Letters}, 109(16), 160401.
\bibitem{yao2017discrete} Yao, N. Y. et al. (2017). Discrete Time Crystals: Rigidity, Criticality, and Realizations. \emph{Physical Review Letters}, 118(3), 030401.
\bibitem{else2016floquet} Else, D. V., Bauer, B., \& Nayak, C. (2016). Floquet Time Crystals. \emph{Physical Review Letters}, 117(9), 090402.
\bibitem{frohlich1968long} Fröhlich, H. (1968). Long-range coherence and energy storage in biological systems. \emph{International Journal of Quantum Chemistry}, 2(5), 641–649.
\bibitem{friston2010free} Friston, K. (2010). The free-energy principle: a unified brain theory? \emph{Nature Reviews Neuroscience}, 11(2), 127–138.
\bibitem{hameroff2014consciousness} Hameroff, S., \& Penrose, R. (2014). Consciousness in the universe: A review of the 'Orch OR' theory. \emph{Physics of Life Reviews}, 11(1), 39–78.
\end{thebibliography}
\end{document}
