% Auto-generated publication draft
\section*{Autodraft: The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not}
\textbf{Thread:} cosmos\\
\textbf{Digest slug:} auto-2026-02-23-the-large-cores-of-dark-matter-and-globular-clusters-in-as1063-p\\
\textbf{Source:} \texttt{site/cosmos/digests/auto-2026-02-23-the-large-cores-of-dark-matter-and-globular-clusters-in-as1063-p.html}\\
\textbf{Generated:} 2026-02-23 14:59 UTC\\

\subsection*{Structured draft body}
Autodraft · The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications About Autodraft: The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not Date: 2026-02-23 · Thread: cosmos · Status: hypothesis · Confidence: low Source chatgpt/pdf/The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf Draft summary Autogenerated stub from corpus scan. Expand with validated claims and citation links before promotion. Key findings (draft) Pending manual extraction. Evidence \& citations TODO: add primary references. Hypotheses TODO: explicitly label speculative claims. Next queries What is the smallest testable claim from this source?

\subsection*{Validation checklist}
\begin{itemize}
  \item Verify all nontrivial claims against the original source.
  \item Add explicit citations/DOIs where available.
  \item Mark confidence for each key claim (low/medium/high).
\end{itemize}
