\section*{Cosmos Time Crystals Holography Minimal Ground Publication V1}
\subsection*{Abstract}
This publication provides a structured synthesis for Cosmos Time Crystals Holography Minimal Ground Publication V1, with claim-to-evidence framing and a validation path for downstream readers.

\subsection*{Keywords}
cosmos, research, publication

\subsection*{Main Content}
% Journal-style manuscript
\title{Time Crystals, Holography, and Quantum Information:\\A Minimal Coherence Framework}
\author{Cohera Lab}
\date{February 2026}
\maketitle

\begin{abstract}
We present a compact synthesis connecting Floquet time-crystal dynamics, holographic coarse-graining, and operator-level coherence constraints. Rather than over-claiming unification, we construct a minimal testable scaffold that separates established results from speculative bridges. The framework defines measurable coherence criteria and explicit falsification hooks suitable for numerical and conceptual evaluation.
\end{abstract}

\paragraph{Keywords} time crystals; Floquet dynamics; holography; quantum information; coherence operators

\section{Introduction}
Time-crystal discourse often blends distinct layers: equilibrium no-go results, non-equilibrium Floquet order, and broader information-theoretic interpretations. This paper imposes a minimal common language so each claim is tied to assumptions, observables, and failure conditions.

\section{Coherence Functional}
Let $\rho_t$ denote a reduced state over a bounded operator algebra. We define a periodic coherence witness over one drive period $T$:
\begin{equation}
\mathcal{C}_T = \frac{1}{T}\int_0^T \operatorname{Tr}(\rho_t\,\mathcal{O}_c)\,dt,
\end{equation}
where $\mathcal{O}_c$ is a coherence-sensitive bounded observable.

For perturbation family $\{\epsilon\}$, coherence drift is
\begin{equation}
\Delta_{\epsilon}=\left|\mathcal{C}_T(\epsilon)-\mathcal{C}_T(0)\right|.
\end{equation}

\section{Stability Threshold}
With stroboscopic map $\Phi_T$, robust regime membership is declared only when
\begin{equation}
\sup_{\epsilon\in[0,\epsilon_0]}\Delta_{\epsilon} < \delta_*,
\end{equation}
for pre-registered tolerance $\delta_*$. Violations indicate fragile coherence and block strong interpretive claims.

\section{Emergent Geometry Protocol}
We compare boundary and reconstructed proxies using a coarse-graining channel $\mathcal{R}$. Consistency is evaluated by
\begin{equation}
\left|\mathcal{C}^{\partial}_T-\mathcal{C}^{\mathrm{bulk}}_T\right| \le \eta,
\end{equation}
where $\eta$ is reported as model error (not absorbed narratively). In local parameter chart $\theta$, geometry diagnostics can be approximated through Fisher information:
\begin{equation}
g_{ab}(\theta)=\sum_x p(x\mid\theta)\,\partial_a\log p(x\mid\theta)\,\partial_b\log p(x\mid\theta).
\end{equation}

\section*{Validation and Falsification}
\begin{itemize}
\item Sweep drive-noise amplitude and detect breakdown of the stability threshold.
\item Replace $\mathcal{O}_c$ with orthogonal witness families to test basis sensitivity.
\item Reject holographic interpretations if boundary--bulk coherence mismatch remains high in controlled toy models.
\end{itemize}

\section*{Conclusion}
The manuscript provides a cleaner publication baseline: assumptions are explicit, equations are legible, and interpretive claims are paired with concrete failure criteria.
