% Auto-generated publication draft
\section*{Autodraft: The large cores of dark matter and globular clusters in AS1063 Possible evidence of self-interacting dark matter Or not}
\textbf{Thread:} cosmos\\
\textbf{Digest slug:} auto-the-large-cores-of-dark-matter-and-globular-clusters-in-as1063-possible-evi\\
\textbf{Source:} \texttt{site/cosmos/digests/auto-the-large-cores-of-dark-matter-and-globular-clusters-in-as1063-possible-evi.html}\\
\textbf{Generated:} 2026-02-23 14:59 UTC\\

\subsection*{Structured draft body}
Autodraft · The large cores of dark matter and globular clusters in AS1063 Possible evidence of self-interacting dark matter Or not · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications About Autodraft: The large cores of dark matter and globular clusters in AS1063 Possible evidence of self-interacting dark matter Or not Date: 2026-02-23 · Thread: cosmos · Status: extracted-draft · Confidence: low-medium Source chatgpt/pdf/The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf DOI: not detected automatically. Auto summary (preview-based) Astronomy \& Astrophysics manuscript no. AA\_main February 19, 2026 ©ESO 2026 The large cores of dark matter and globular clusters in AS1063. Possible evidence of self-interacting dark matter. Or not. J.M. Diego1,⋆ Instituto de Física de Cantabria (CSIC-UC). Avda. Los Castros s/n. 39005 Santander, Spain arXiv:2602.15940v1 [astro-ph.GA] 17 Feb 2026 February 19, 2026 ABSTRACT Deep JWST images of AS1063 reveals tens of th Key findings (auto-extracted) Primary topic appears to center on: large, cores, dark, matter. Source was auto-indexed and text-previewed for rapid triage. Needs manual verification before promoting any strong claim to high confidence. Evidence \& citations Source file: chatgpt/pdf/The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf Extraction scope: 1-2 Abstract/preview extracted automatically. Claim → evidence mapping (auto) Claim: AA\_main February 19, 2026 ©ESO 2026 The large cores of dark matter and globular clusters in AS1063. Evidence quote: “AA\_main February 19, 2026 ©ESO 2026 The large cores of dark matter and globular clusters in AS1063.” Page hint: 1-2 Claim: 39005 Santander, Spain arXiv:2602.15940v1 [astro-ph.GA] 17 Feb 2026 February 19, 2026 ABSTRACT Deep JWST images of AS1063 reveals tens of thousands of globular clusters in the gala… Evidence quote: “39005 Santander, Spain arXiv:2602.15940v1 [astro-ph.GA] 17 Feb 2026 February 19, 2026 ABSTRACT Deep JWST images of AS1063 reveals tens of thousands of globular clusters in the galaxy cluster AS1063.” Page hint: 1-2 Claim: When compared with the lensing model based on the same JWST data, the distribution of globular clusters traces closely the distribution of lensing mass (mostly composed of dark mat… Evidence quote: “When compared with the lensing model based on the same JWST data, the distribution of globular clusters traces closely the distribution of lensing mass (mostly composed of dark matter).” Page hint: 1-2 Falsification / validation checklist Re-read full source and verify the central claim sentence-by-sentence. Cross-check against at least one independent source before promotion. Keep confidence at low-medium until replication or corroboration is explicit. Next queries What is the smallest testable claim from this source?

\subsection*{Validation checklist}
\begin{itemize}
  \item Verify all nontrivial claims against the original source.
  \item Add explicit citations/DOIs where available.
  \item Mark confidence for each key claim (low/medium/high).
\end{itemize}
