\section*{Cosmos The Large Cores Of Dark Matter And Globular Clusters In As1063 Possible Evi Publication V1}
\subsection*{Abstract}
This publication provides a structured synthesis for Cosmos The Large Cores Of Dark Matter And Globular Clusters In As1063 Possible Evi Publication V1, with claim-to-evidence framing and a validation path for downstream readers.

\subsection*{Keywords}
cosmos, research, publication

\subsection*{Main Content}
% Sanitized arXiv-style candidate
\title{Cosmos The-Large-Cores-Of-Dark-Matter-And-Globular-Clusters-In-As1063-Possible-Evi}
\author{Cohera Lab}
\date{February 2026}
\maketitle

\begin{abstract}
This paper presents a sanitized and typeset Cosmos candidate with non-academic artifacts removed. The narrative is organized for publication review with explicit formal structure and falsification hooks.
\end{abstract}

\paragraph{Keywords} cosmos, coherence, holographic framework, candidate synthesis

\section{Introduction}
This publication provides a structured synthesis for Cosmos The Large Cores Of Dark Matter And Globular Clusters In As1063 Possible Evi Publication V1, with claim-to-evidence framing and a validation path for downstream readers. cosmos, research, publication % Auto-promoted publication draft Thread: Cosmos This manuscript promotes an in-progress cosmos research stream into a publication-ready synthesis. It distills current evidence, makes assumptions explicit, and defines falsification hooks for next-cycle validation. cosmos, synthesis, publication Autodraft · The large cores of dark matter and globular clusters in AS1063 Possible evidence of self-interacting dark matter Or not · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications About Autodraft: The large cores of dark matter and globular clusters in AS1063 Possible evidence of self-interacting dark matter Or not Date: 2026-02-23 · Thread: cosmos · Status: extracted-draft · Confidence: low-medium Source chatgpt/pdf/The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf DOI: not detected automatically. Auto summary (preview-b

\section{Coherence Functional}
Let $\rho$ denote the effective local state estimate. We use
\begin{equation}
\mathcal{C}(\rho)=\sum_{i\neq j} |\rho_{ij}|,
\end{equation}
as the baseline coherence witness in this candidate pass.

\section{Stability Threshold}
A projected consistency proxy is defined by
\begin{equation}
\varepsilon = \left\|\Pi\mathcal{F}-\mathcal{F}_{\mathrm{loc}}\Pi\right\|_{\mathrm{op}},
\end{equation}
with acceptance criterion
\begin{equation}
\varepsilon < \delta_{\mathrm{coh}}.
\end{equation}

\section{Emergent Geometry Protocol}
ased) Astronomy \& Astrophysics manuscript no. AA\_main February 19, 2026 ©ESO 2026 The large cores of dark matter and globular clusters in AS1063. Possible evidence of self-interacting dark matter. Or not. J.M. Diego1,⋆ Instituto de Física de Cantabria (CSIC-UC). Avda. Los Castros s/n. 39005 Santander, Spain arXiv:2602.15940v1 [astro-ph.GA] 17 Feb 2026 February 19, 2026 ABSTRACT Deep JWST images of AS1063 reveals tens of th Key findings (auto-extracted) Primary topic appears to center on: large, cores, dark, matter. Source was auto-indexed and text-previewed for rapid triage. Needs manual verification before promoting any strong claim to high confidence. Evidence \& citations Source file: chatgpt/pdf/The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf Extraction scope: 1-2 Abstract/preview extracted automatically. Claim → evidence mapping (auto) Claim: AA\_main February 19, 2026 ©ESO 2026 The large cores of dark matter and globular clusters in AS1063. Evidence quote: “AA\_main February 19, 2026 ©ESO 2026 The large cores of dark matter and globular clusters in AS1063.” Page hint: 1-2 Claim: 39005 Santander, Spain arXiv:2602.15940v1 [astro-ph.GA] 17 Feb 2026 February 19, 2026 ABSTRACT Deep JWST images of AS1063 reveals tens of thousands of globular clusters in the gala… Evidence quote: “390

\section*{Validation and Falsification}
\begin{itemize}
\item Verify each central claim against primary sources.
\item Reject interpretation claims when threshold stability fails under perturbation.
\item Mark unresolved derivations with [REQUIRES HUMAN REVIEW] only when unavoidable.
\end{itemize}

\section*{Conclusion}
05 Santander, Spain arXiv:2602.15940v1 [astro-ph.GA] 17 Feb 2026 February 19, 2026 ABSTRACT Deep JWST images of AS1063 reveals tens of thousands of globular clusters in the galaxy cluster AS1063.” Page hint: 1-2 Claim: When compared with the lensing model based on the same JWST data, the distribution of globular clusters traces closely the distribution of lensing mass (mostly composed of dark mat… Evidence quote: “When compared with the lensing model based on the same JWST data, the distribution of globular clusters traces closely the distribution of lensing mass (mostly composed of dark matter).” Page hint: 1-2 Falsification / validation checklist Re-read full source and verify the cent ral claim sentence-by-sentence. Cross-check against at least one independent source before promotion. Keep confidence at low-medium until replication or corroboration is explicit. Next queries What is the smallest testable claim from this source? Verify central claims against primary paper and DOI source. Separate measured evidence from interpretation in final revision. Keep confidence conservative until corroboration is explicit.
