\section*{Cosmos The Large Cores Of Dark Matter And Globular Clusters In As1063 Possible Evidence Of Self I Publication V1}
\subsection*{Abstract}
This publication provides a structured synthesis for Cosmos The Large Cores Of Dark Matter And Globular Clusters In As1063 Possible Evidence Of Self I Publication V1, with claim-to-evidence framing and a validation path for downstream readers.

\subsection*{Keywords}
cosmos, research, publication

\subsection*{Main Content}
% Auto-promoted publication draft
\title{The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not}
\author{Cohera Lab}
\date{2026-02-23}
\maketitle

\begin{center}\small Thread: Cosmos\end{center}

\begin{abstract}
This manuscript promotes an in-progress cosmos research stream into a publication-ready synthesis. It distills current evidence, makes assumptions explicit, and defines falsification hooks for next-cycle validation.
\end{abstract}

\paragraph{Keywords} cosmos, synthesis, publication

\section{Introduction}
Autodraft · The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not Cohera Lab Home Research Publications Autodraft: The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not Thread: cosmos · Date: 2026-02-23 · [STATUS: READY FOR SYNTHESIS] Source The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf Core thesis and falsification hook extracted in pipeline pass.

\section{Evidence and Claims}
Evidence extraction pending deeper review.

\section{Discussion}
Discussion to be expanded in subsequent research cycles.

\section{Validation Hooks}
\begin{itemize}
\item Verify central claims against primary paper and DOI source.
\item Separate measured evidence from interpretation in final revision.
\item Keep confidence conservative until corroboration is explicit.
\end{itemize}
