\section*{Cosmos 2026 02 23 The Large Cores Of Dark Matter And Globular Clusters In As1063 P Publication V1}
\subsection*{Abstract}
This publication provides a structured synthesis for Cosmos 2026 02 23 The Large Cores Of Dark Matter And Globular Clusters In As1063 P Publication V1, with claim-to-evidence framing and a validation path for downstream readers.

\subsection*{Keywords}
cosmos, research, publication

\subsection*{Main Content}
% Sanitized arXiv-style candidate
\title{Cosmos 2026-02-23-The-Large-Cores-Of-Dark-Matter-And-Globular-Clusters-In-As1063-P}
\author{Cohera Lab}
\date{February 2026}
\maketitle

\begin{abstract}
This paper presents a sanitized and typeset Cosmos candidate with non-academic artifacts removed. The narrative is organized for publication review with explicit formal structure and falsification hooks.
\end{abstract}

\paragraph{Keywords} cosmos, coherence, holographic framework, candidate synthesis

\section{Introduction}
This publication provides a structured synthesis for Cosmos 2026 02 23 The Large Cores Of Dark Matter And Globular Clusters In As1063 P Publication V1, with claim-to-evidence framing and a validation path for downstream readers. cosmos, research, publication % Auto-promoted publication draft Thread: Cosmos This manuscript promotes an in-progress cosmos research stream into a publication-ready synthesis. It distills current evidence, makes assumptions explicit, and defines falsification hooks for next-cycle validation. cosmos, synthesis, publication Autodraft · The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications About Autodraft: The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not Date: 2026-02-23 · Thread: cosmos · Status: hypothesis · Confidence: low Source chatgpt/pdf/The\_large\_cores\_of\_dark\_matter\_and\_globular\_clusters\_in\_AS1063\_Possible\_evidence\_of\_self-interacting\_dark\_matter\_Or\_not.pdf Draft summary Autogenerated stub

\section{Coherence Functional}
Let $\rho$ denote the effective local state estimate. We use
\begin{equation}
\mathcal{C}(\rho)=\sum_{i\neq j} |\rho_{ij}|,
\end{equation}
as the baseline coherence witness in this candidate pass.

\section{Stability Threshold}
A projected consistency proxy is defined by
\begin{equation}
\varepsilon = \left\|\Pi\mathcal{F}-\mathcal{F}_{\mathrm{loc}}\Pi\right\|_{\mathrm{op}},
\end{equation}
with acceptance criterion
\begin{equation}
\varepsilon < \delta_{\mathrm{coh}}.
\end{equation}

\section{Emergent Geometry Protocol}
 from corpus scan. Expand with validated claims and citation links before promotion. Key findings (draft) Pending manual extraction. Evidence \& citations TODO: add primary references. Hypotheses TODO: explicitly label speculative claims. Next queries What is the smallest testable claim from this source? Evidence extraction pending deeper review. Discussion to be expanded in subsequent research cycles. Verify central claims against primary paper and DOI source. Separate measured evidence from interpretation in final revision. Keep confidence conservative until corroboration is explicit.

\section*{Validation and Falsification}
\begin{itemize}
\item Verify each central claim against primary sources.
\item Reject interpretation claims when threshold stability fails under perturbation.
\item Mark unresolved derivations with [REQUIRES HUMAN REVIEW] only when unavoidable.
\end{itemize}

\section*{Conclusion}
Discussion is constrained to explicit claims and falsification hooks.
