\section*{Ethos 2026 02 23 How Should Ai Knowledge Be Governed Epistemic Authority Structur Publication V1}
\subsection*{Abstract}
This publication provides a structured synthesis for Ethos 2026 02 23 How Should Ai Knowledge Be Governed Epistemic Authority Structur Publication V1, with claim-to-evidence framing and a validation path for downstream readers.

\subsection*{Keywords}
ethos, research, publication

\subsection*{Main Content}
% Auto-promoted publication draft
\title{How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs}
\author{Cohera Lab}
\date{2026-02-23}
\maketitle

\begin{center}\small Thread: Ethos\end{center}

\begin{abstract}
This manuscript promotes an in-progress ethos research stream into a publication-ready synthesis. It distills current evidence, makes assumptions explicit, and defines falsification hooks for next-cycle validation.
\end{abstract}

\paragraph{Keywords} ethos, synthesis, publication

\section{Introduction}
Autodraft · How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications About Autodraft: How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs Date: 2026-02-23 · Thread: ethos · Status: hypothesis · Confidence: low Source chatgpt/pdf/How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs.pdf Draft summary Autogenerated stub from corpus scan. Expand with validated claims and citation links before promotion. Key findings (draft) Pending manual extraction. Evidence \& citations TODO: add primary references. Hypotheses TODO: explicitly label speculative claims. Next queries What is the smallest testable claim from this source?

\section{Evidence and Claims}
Evidence extraction pending deeper review.

\section{Discussion}
Discussion to be expanded in subsequent research cycles.

\section{Validation Hooks}
\begin{itemize}
\item Verify central claims against primary paper and DOI source.
\item Separate measured evidence from interpretation in final revision.
\item Keep confidence conservative until corroboration is explicit.
\end{itemize}
