\section*{Research Literacy Quickstart + Confidence Rubric}
\subsection*{Abstract}
This publication provides a structured synthesis for Research Literacy Quickstart + Confidence Rubric, with claim-to-evidence framing and a validation path for downstream readers.

\subsection*{Keywords}
ethos, research, publication

\subsection*{Main Content}
% Auto-generated publication draft
\section*{Research Literacy Quickstart + Confidence Rubric}
\textbf{Thread:} ethos\\
\textbf{Digest slug:} research-literacy-confidence-rubric\\
\textbf{Source:} \texttt{site/ethos/digests/research-literacy-confidence-rubric.html}\\
\textbf{Generated:} 2026-02-23 14:59 UTC\\

\subsection*{Structured draft body}
Research Literacy and Confidence Rubric · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications Contact About Research Literacy Quickstart + Confidence Rubric Date: 2026-02-21 · Tags: ethos, epistemology, education, confidence · Confidence: high Summary This module defines a simple quality filter for all Cohera threads. It separates claims by evidence strength, marks uncertainty explicitly, and provides a repeatable paper-reading workflow. Quick paper-reading workflow Read abstract + conclusion first: what was actually tested? Check methods and sample constraints. Distinguish correlation vs intervention. Check if result has replication/systematic review support. Write one-sentence claim with confidence label. Confidence rubric (standard) High: converging evidence from reviews/meta-analyses or strong replicated primary data. Medium: plausible and supported by limited or mixed evidence. Low: early, speculative, or mostly conceptual; publish only as hypothesis. Common traps Single-paper overreach Mechanism inflation beyond data Confusing surrogate markers with outcomes Ignoring population/context limits Citations Guyatt et al. (GRADE framework) — https://www.bmj.com/content/336/7650/924 Ioannidis (Why Most Published Research Findings Are False) — https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.0020124 Cochrane Handbook (evidence synthesis standards) — https://training.cochrane.org/handbook Next iteration Add a per-thread checklist card so every digest visibly includes: confidence, citations, uncertainty, and known failure modes. Ethos module v1 · show your work, mark uncertainty.

\subsection*{Validation checklist}
\begin{itemize}
  \item Verify all nontrivial claims against the original source.
  \item Add explicit citations/DOIs where available.
  \item Mark confidence for each key claim (low/medium/high).
\end{itemize}
