% Auto-generated publication draft
\section*{Autodraft: How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs}
\textbf{Thread:} ethos\\
\textbf{Digest slug:} auto-2026-02-23-how-should-ai-knowledge-be-governed-epistemic-authority-structur\\
\textbf{Source:} \texttt{site/ethos/digests/auto-2026-02-23-how-should-ai-knowledge-be-governed-epistemic-authority-structur.html}\\
\textbf{Generated:} 2026-02-23 14:59 UTC\\

\subsection*{Structured draft body}
Autodraft · How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs · Cohera Lab Cohera Lab Home Research Cosmos Regenesis Ethos Publications About Autodraft: How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs Date: 2026-02-23 · Thread: ethos · Status: hypothesis · Confidence: low Source chatgpt/pdf/How\_should\_AI\_knowledge\_be\_governed\_Epistemic\_authority\_structural\_transparency\_and\_the\_case\_for\_open\_cognitive\_graphs.pdf Draft summary Autogenerated stub from corpus scan. Expand with validated claims and citation links before promotion. Key findings (draft) Pending manual extraction. Evidence \& citations TODO: add primary references. Hypotheses TODO: explicitly label speculative claims. Next queries What is the smallest testable claim from this source?

\subsection*{Validation checklist}
\begin{itemize}
  \item Verify all nontrivial claims against the original source.
  \item Add explicit citations/DOIs where available.
  \item Mark confidence for each key claim (low/medium/high).
\end{itemize}
